### 2021-03-04 Rmd/Report cleaning.Rmd
Start a Rmd and more updates

---
title: "Flatiron report"
author: "Christelle Colin-Leitzinger"
date: `r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    # keep_md: true
---

<style type="text/css">
.figure {
   margin-top: 25px;
   <!-- margin-bottom: 100px; -->
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

# Report
<br>
<br>

***

```{r library, include=FALSE}
library(drake)
```

```{r}
loadd(clinical_data, vitals, drugs, auc, creatinine)
```

## Creatinine cleaning
.  
```{r creatinine}

```