### 2021-03-04 Rmd/Report cleaning.Rmd
Start report

---
title: "Flatiron data cleaning and analysis report"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: flatly
    highlight: pygments
    df_print: paged
---

<style type="text/css">
.figure {
   margin-top: 25px;
   <!-- margin-bottom: 100px; -->
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

# Data cleaning
***
<br>

```{r library, include=FALSE}
library(drake)
library(tidyverse)
library(data.table)
library(lubridate)
library(ggplot2)
library(gtsummary)
library(survminer)
library(survival)
```

```{r loadd}
loadd(clinical_data, vitals, drugs, auc, creatinine)
```

## Creatinine cleaning

Only creatinine from blood and serum are being used. All units are correct (mg/dL).  
Here an sample of the non-cleaned data with really low `testresultcleaned`.

```{r creatinine, rows.print=15, max.print=15}
creatinine %>% filter(str_detect(labcomponent, "blood|serum")) %>% 
arrange(testresultcleaned)
```
All `testresultcleaned` and `testresult` variable are cleaned as :  
-creat < 0.8 is set at 0.8  
-creat > 3 are eliminated  
To create the final `creatinine` variable ,`testresultcleaned` and `testresult` are both used but `testresultcleaned` is used over `testresult` when both are available.

Here a sample of the cleaned data
```{r creatinine1, rows.print=15, max.print=15}
creatinine <- creatinine %>%
  # remove creatine urine
  filter(str_detect(labcomponent, "blood|serum")) %>% 
  mutate(creatinine = case_when(
    testresultcleaned < 0.8        ~ 0.8,
    testresultcleaned > 3          ~ NA_real_,
    TRUE                           ~ testresultcleaned
  )) %>% 
  # F092F8B827DDC kept umol/L as they actually are mg/dL range 45 to 90 μmol/L (0.5 to 1.0 mg/dL) for women.
  mutate(creatinine1 = case_when(
    testresult < 0.8               ~ 0.8,
    testresult > 3                 ~ NA_real_,
    TRUE                           ~ testresult
  )) %>% 
  mutate(creatinine = coalesce(creatinine, creatinine1)) %>% 
  filter(!is.na(creatinine)) %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y")) %>% 
  select(c("patientid", creat_date = "testdate", "creatinine", creat_units = "testunitscleaned"))
creatinine %>% distinct(patientid, .keep_all = TRUE)

vitals <- vitals %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y"))
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)")
```
<br>

## Body surface area cleaning

`testresultcleaned` is all within expected range but only few data is available (`r NROW(body_surface_area %>% filter(!is.na(testresultcleaned)) %>% distinct(patientid, .keep.all=TRUE))` patients). Even less with an associated date!

```{r bsa}
body_surface_area
```

By using `testresult` and before cleaning only `r NROW(body_surface_area %>% filter(!is.na(body_surface_area$testresult)) %>% distinct(patientid, .keep.all=TRUE))` patients have data for their BSA. So I will also use the height and weight of patients to calculate their BSA by using the Du Bois formula.

To create the final `BSA` variable ,`testresultcleaned` is used over `testresult` when both are available.  
`testresult` was further cleaned as data < 0.418 and > 4.14 are eliminated. These values were calculated from the biggest and the smallest individuals recorded.  
I also remove the outliers within patient (2 examples here).

```{r}
vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)" & patientid == "F2D5A8748B268") %>% 
  select(patientid, testdate, testresult, testresultcleaned)
vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)" & patientid == "F875EA1B044F0") %>% 
  select(patientid, testdate, testresult, testresultcleaned)
# F3AE4546B5D12
```

Here a sample of the cleaned data

```{r bsa1, rows.print=15, max.print=15}
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)") %>% # only 399 patients data
  # Limit max = 4.14, min = 0.418 calculated for biggest to smallest person
  mutate(testresult = ifelse(testresult < 0.418 | testresult > 4.14, NA_real_, testresult)) %>% 
  group_by(patientid) %>%
  # I used the median to remove outlier within a patient
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(BSA = case_when(
    testresult > (median_testresult + 0.4) |
      testresult < (median_testresult - 0.4)      ~ NA_real_,
    TRUE                                          ~ testresult
  )) %>% 
  mutate(BSA = coalesce(testresultcleaned, BSA)) %>% 
  filter(!is.na(BSA)) %>% 
  mutate(BSA_units = "m2") %>% 
  select(c("patientid", bsa_date = "testdate", "BSA", "BSA_units"))

body_surface_area %>% distinct(patientid, .keep_all = TRUE)
```
<br>

## Height

Height data complicated. Again you have `testresultcleaned` and `testresult` with `testresultcleaned` which will prioritized.  
`testresultcleaned` has units in cm but some data are wrong. `testresult` has results in cm, ft or in. It is pretty easy to clean data between these 3 units because the measurement cannot overlap.
Both are cleaned as < 120 or > 231 are eliminated. These value are for the shortest and tallest women recorded.
The units can be filled by patients when absent.

```{r height}
vitals %>% 
  filter(test == "body height" & patientid == "FCB6DAAFF6527")
# FC109A6F719E7
# FC106805F9AB9
# F91590CB83EC8 weird
```

Outliers are eliminated as much as possible.

```{r height1}
vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  filter(patientid == "F01413C06D921") %>% 
  select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified,
           testunitscleaned, testunits, testresult))
```

PB: Some patient don't have dates but have data. Should I take a mean of all the height measured or the first one - If I have to choose I would take the mean (patients shouldn't grow or shorten much)?
If all dates are NA for a patients, I cannot combined `height` with `episodedate` of the drug data.

Here a samples of the data.
```{r height2, rows.print=15, max.print=15}
height <- vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  # select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified, 
  #          testunitscleaned, testunits, testresult)) %>%

  group_by(patientid, testunits) %>%
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(testresult_verified = case_when(
    testresult > (median_testresult + 3) & 
      testunits == "in" |
      testresult < (median_testresult - 3) & 
      testunits == "in"                         ~ NA_real_,
    testresult > (median_testresult + 5) & 
      testunits == "cm" |
      testresult < (median_testresult - 5) & 
      testunits == "cm"                         ~ NA_real_,
    TRUE                                        ~ testresult
  )) %>% 
  # select(c("patientid", testdate, testresult, median_testresult, testresult_verified,
  #          testunits, testunitscleaned, testresultcleaned, testresultcleaned_verified))

  mutate(height = case_when(
    !is.na(testresultcleaned_verified)          ~ round(testresultcleaned_verified / 100, 3), # cm to m
    testunits == "cm"                           ~ round(testresult_verified / 100, 3),
    testunits == "ft"                           ~ round(testresult_verified / 3.281, 3), # ft to m
    testunits == "in" | 
      !is.na(testresult)                        ~ round(testresult_verified / 39.37, 3) # in to m
    )) %>% 
  filter(!is.na(height)) %>% 
  mutate(height_units = "m") %>% 
  select(c("patientid", height_date = "testdate", "height", "height_units"))

height %>% distinct(patientid, .keep_all = TRUE)
```
`r emo::ji("check")`

<br>

## Weight

The more complex cleaning (before the drugs...`r emo::ji("sad")`). Contrary to height with in cm, ft or in which don't overlap. Weight has data in kg.......

```{r weight}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  ))
```

## Drugs

`r emo::ji("face_screaming_in_fear")`