### 2021-03-19 Rmd/Report analysis 03172021.Rmd
Finished Rmd prelim results

---
title: "Flatiron analysis"
author: "Christelle Colin-Leitzinger"
date: 'r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: flatly
    highlight: pygments
    df_print: paged
---

<style type="text/css">
.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 1px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

# Analysis
***
<br>
For all these analysis, I excluded the patients Lauren said to remove from the analysis.

```{r library, include=FALSE}
library(drake)
library(tidyverse)
library(data.table)
library(lubridate)
library(ggplot2)
library(gtsummary)
library(survminer)
library(survival)
```

```{r loadd}
loadd(clinical_data, vitals, drugs, auc, creatinine)
Frontline <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/Frontline.rds")
```

## Clinical summary table
```{r analysis_data}
analysis_data <- Frontline %>% 
  distinct(patientid, chemotherapy_type, .keep_all = TRUE) %>% 
  filter(exclude != "1")
```

<span style="color:green">**Table 1 Overall summary demographic/clinical data**</span>
```{r Table 1}
analysis_data %>% 
  select(c(race, raceeth, ethnicity, "vital",
           agecat, "ageatdx",

           "bmi", bmi_cat, "BSA",

           "histology", "groupstage", "tstage", "stagecat",

           "brca1_germline", "brca2_germline", "brca_germline", "brca1_tumor",
           "brca2_tumor", "brca_tumor", "brca1_unknown", "brca2_unknown", "brca_unknown",

           "hrd_tissue", "gis_tissue")) %>% 
  tbl_summary() %>% bold_labels()
```

<span style="color:green">**Table 2 Summary demographic/clinical data by race**</span>
```{r Table 2}
analysis_data %>% 
  select(c(race, raceeth, ethnicity, "vital",
           agecat, "ageatdx",

           "bmi", bmi_cat, "BSA",

           "histology", "groupstage", "tstage", "stagecat",

           "brca1_germline", "brca2_germline", "brca_germline", "brca1_tumor",
           "brca2_tumor", "brca_tumor", "brca1_unknown", "brca2_unknown", "brca_unknown",

           "hrd_tissue", "gis_tissue")) %>% 
  tbl_summary(by = raceeth) %>% bold_labels() %>% add_overall() %>% add_p()
```
<br>

## Treatment summary table
<span style="color:green">**Table 3 Summary treatment data by race**</span>

```{r Table 3}
tbl1 <- analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
  filter(exclude != "1") %>% 
  select(c(raceeth, delay_incare_dx_to_treat, #mean_skipped_cycle_indays,
           "issurgery", "extentofdebulking", "resdz", "debulking", "diff_surg_dx",
           "chemotherapy_type", "therapy", "mean_RDI_per_drug_chemotype", "RDI_grp")) %>% 
  tbl_summary(by = raceeth) %>% bold_labels() %>% add_overall() %>% add_p()
tbl2 <- analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
  filter(exclude != "1") %>% 
  select(c(raceeth, delay_incare_dx_to_treat, #mean_skipped_cycle_indays,
           "issurgery", "extentofdebulking", "resdz", "debulking", "diff_surg_dx",
           "chemotherapy_type", "therapy", "treatment_sequence", "mean_RDI_per_drug_chemotype", "RDI_grp")) %>% 
  tbl_summary(by = raceeth, missing = "no") %>% bold_labels() %>% add_overall() %>% add_p()
tbl_merge(list(tbl1, tbl2), tab_spanner = c("with unknown", "without unknown"))
```

<span style="color:green">**Table 4 Summary RDI data by race and bmi**</span>

```{r Table 4}
# Treatment & demo
tbl1 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(raceeth,  "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = raceeth,
              label = list(mean_RDI_per_drug_chemotype ~ "carb/adjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl2 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(bmi_cat,  "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = bmi_cat,
              label = list(mean_RDI_per_drug_chemotype ~ "carb/adjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl3 <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))

tbl1 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = raceeth,
              label = list(mean_RDI_per_drug_chemotype ~ "carb/neoadjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl2 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = bmi_cat,
              label = list(mean_RDI_per_drug_chemotype ~ "carb/neoadjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl4 <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))

tbl1 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Adjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = raceeth,
              label = list(mean_RDI_per_drug_chemotype ~ "taxel/adjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl2 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Adjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = bmi_cat,
              label = list(mean_RDI_per_drug_chemotype ~ "taxel/adjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl5 <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))

tbl1 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Neoadjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = raceeth,
              label = list(mean_RDI_per_drug_chemotype ~ "taxel/neoadjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl2 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Neoadjuvant") %>% 
  filter(exclude != "1") %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
  tbl_summary(by = bmi_cat,
              label = list(mean_RDI_per_drug_chemotype ~ "taxel/neoadjuvant \nRDI")) %>%
  bold_labels() %>% add_p()
tbl6 <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))

tbl_stack(list(tbl3, tbl4, tbl5, tbl6))
```
<br>

## Survivals
### OS by clinical

<span style="color:green">**Survival plots from date of diagnosis on the left and from date of treatment on the left**</span>

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
# OS by race
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ raceeth, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by raceeth
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ raceeth, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by histology
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ histology, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by tstage
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ tstage, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by bmi_cat
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ bmi_cat, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca1_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca1_tumor, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca2_tumor, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca1_germline, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca2_germline, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
# OS by race
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ raceeth, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by raceeth
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ raceeth, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by histology
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ histology, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by tstage
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ tstage, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by bmi_cat
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ bmi_cat, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca1_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca1_tumor, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca2_tumor, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca1_germline, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca2_germline, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))
```
</div>
</div>

<br>

### OS by treatment

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ issurgery, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by extentofdebulking
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ extentofdebulking, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by residualdiseasestatus
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ residualdiseasestatus, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by chemotherapy_type
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ chemotherapy_type, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by treatment_sequence
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ treatment_sequence, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by therapy
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ therapy, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp for platin
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     filter((exclude != "1") & str_detect(drugname, "platin")) %>%

                     distinct(patientid, .keep_all = TRUE)), 
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp for taxel
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     filter((exclude != "1") & str_detect(drugname, "taxel")) %>%

                     distinct(patientid, .keep_all = TRUE)), 
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ issurgery, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by extentofdebulking
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ extentofdebulking, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by residualdiseasestatus
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ residualdiseasestatus, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by chemotherapy_type
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ chemotherapy_type, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by treatment_sequence
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ treatment_sequence, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by therapy
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ therapy, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp for platin
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     filter((exclude != "1") & str_detect(drugname, "platin")) %>%

                     distinct(patientid, .keep_all = TRUE)), 
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))

# OS by RDI_grp for taxel
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
                     filter((exclude != "1") & str_detect(drugname, "taxel")) %>%

                     distinct(patientid, .keep_all = TRUE)), 
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 4.5, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 2))
```
</div>
</div>

<br>

## Hazard Ratio

<span style="color:green">**Table 5 HR**</span>  
`coxph(Surv(time = month_at_os, event = vitalstatus)~variables`
```{r}
# tbl1 <- analysis_data %>% 
#   select(bmi_cat) %>% 
#   tbl_summary(#by = vitalstatus,
#               statistic = all_categorical() ~ "{analysis_data %>% filter(vital == 'Dead') %>% NROW()} death ({n})"
#               )
# tbl2 <- coxph(Surv(time = month_at_os,
#                      event = vitalstatus)~bmi_cat,
#                 data = analysis_data) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# 
# tbl_merge(list(tbl1, tbl2))
# 
# coxph(Surv(time = month_at_os, 
#                      event = vitalstatus)~RDI_grp, 
#                 data = analysis_data) %>% 
#   tbl_uvregression(y = vitalstatus, exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()

analysis_data %>% select(vitalstatus, bmi, bmi_cat, raceeth) %>% 
tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = analysis_data$month_at_os, event = analysis_data$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels() #%>%  modify_spanning_header(all_stat_cols() ~ "**Treatment Received**")

tbl1 <- analysis_data %>% 
  filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant")
tbl1 <- tbl1 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl1$month_at_os, event = tbl1$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl2 <- analysis_data %>% 
  filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant")
tbl2 <- tbl2 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl2$month_at_os, event = tbl2$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl3 <- analysis_data %>% 
  filter(drugname == "carboplatin" & str_detect(chemotherapy_type, "Chemo"))
tbl3 <- tbl3 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl3$month_at_os, event = tbl3$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl_stack(list(tbl1, tbl2, tbl3),
          group_header = c("carboplatin/Neoadjuvant", "carboplatin/Adjuvant", "carboplatin/Chemo only")) %>% 
  bold_labels() %>%
  italicize_levels()
# Platin
tbl4 <- analysis_data %>% 
  filter(str_detect(drugname,"platin") & chemotherapy_type == "Neoadjuvant")
tbl4 <- tbl4 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl4$month_at_os, event = tbl4$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl5 <- analysis_data %>% 
  filter(str_detect(drugname,"platin") & chemotherapy_type == "Adjuvant")
tbl5 <- tbl5 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl5$month_at_os, event = tbl5$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl6 <- analysis_data %>% 
  filter(str_detect(drugname,"platin") & str_detect(chemotherapy_type, "Chemo"))
tbl6 <- tbl6 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl6$month_at_os, event = tbl6$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl_stack(list(tbl4, tbl5, tbl6),
          group_header = c("platin/Neoadjuvant", "platin/Adjuvant", "platin/Chemo only")) %>% 
  bold_labels() %>%
  italicize_levels()

tbl7 <- analysis_data %>% 
  filter(str_detect(drugname,"paclitaxel") & chemotherapy_type == "Neoadjuvant")
tbl7 <- tbl7 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl7$month_at_os, event = tbl7$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl8 <- analysis_data %>% 
  filter(str_detect(drugname,"paclitaxel") & chemotherapy_type == "Adjuvant")
tbl8 <- tbl8 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% 
  tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl8$month_at_os, event = tbl8$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  modify_header(update = list(label ~ "**paclitaxel/Adjuvant**"))
tbl9 <- analysis_data %>% 
  filter(str_detect(drugname,"paclitaxel") & str_detect(chemotherapy_type, "Chemo"))
tbl9 <- tbl9 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl9$month_at_os, event = tbl9$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl_stack(list(tbl7, tbl8, tbl9),
          group_header = c("paclitaxel/Neoadjuvant", "paclitaxel/Adjuvant", "paclitaxel/Chemo only")) %>% 
  bold_labels() %>%
  italicize_levels()

tbl10 <- analysis_data %>% 
  filter(str_detect(drugname, "paclitaxel") & chemotherapy_type == "Neoadjuvant")
tbl10 <- tbl10 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl10$month_at_os, event = tbl10$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl11 <- analysis_data %>% 
  filter(str_detect(drugname, "paclitaxel") & chemotherapy_type == "Adjuvant")
tbl11 <- tbl11 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% 
  tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl11$month_at_os, event = tbl11$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  modify_header(update = list(label ~ "**paclitaxel/Adjuvant**"))
tbl12 <- analysis_data %>% 
  filter(str_detect(drugname,"taxel") & str_detect(chemotherapy_type, "Chemo"))
tbl12 <- tbl12 %>% 
  select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
                 y = (Surv(time = tbl12$month_at_os, event = tbl12$vitalstatus)),
                 exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
tbl_stack(list(tbl10, tbl11, tbl12),
          group_header = c("taxel/Neoadjuvant", "taxel/Adjuvant", "taxel/Chemo only")) %>% 
  bold_labels() %>%
  italicize_levels()
```

## Regression
### By clinical, demo
`glm(vitalstatus ~ bmi + bmi_cat + RDI_grp + raceeth, data = analysis_data, family = binomial)`
```{r}
# Regression vital status prediction
model <- glm(vitalstatus ~ bmi + bmi_cat + RDI_grp + raceeth, data = analysis_data, family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

`glm(vitalstatus ~ histology + tstage + brca1_tumor + brca2_tumor + issurgery + chemotherapy_type, data = analysis_data, family = binomial)`
```{r}
model <- glm(vitalstatus ~ 
               histology + tstage + brca1_tumor + brca2_tumor + 
             issurgery +
             chemotherapy_type, 
             data = analysis_data, family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

`glm(RDI_grp ~ agecat + bmi_cat + raceeth + stagecat + histology + chemotherapy_type, data = analysis_data, family = binomial)`
```{r}
# Regression RDI prediction
model <- glm(RDI_grp ~ agecat + bmi_cat + raceeth + stagecat + histology + chemotherapy_type, data = analysis_data, family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

### By drug
```{r}
#  for carbo
model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
               filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant"), family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl3 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))

model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
               filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant"), family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl4 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))

# tbl5 <- 
  tbl_merge(list(tbl3, tbl4), tab_spanner = c("**carboplatin/Adjuvant**", "**carboplatin/Neoadjuvant**"))

#  for taxel
model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
               filter(str_detect(drugname,"taxel") & (chemotherapy_type == "Adjuvant")), family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl3 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))

model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
               filter(str_detect(drugname,"taxel") & (chemotherapy_type == "Neoadjuvant")), family = binomial)
tbl1 <- tbl_regression(model)
tbl2 <- tbl_regression(model, exponentiate = TRUE)
tbl4 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))

# tbl6 <- 
  tbl_merge(list(tbl4, tbl3), tab_spanner = c("**taxel/Adjuvant**", "**taxel/Neoadjuvant**"))
```

## Plots for first treatment
For RDI, it is calculated for neoadjuvant and adjuvant when a patient got both.
<div class = "row">
<div class = "col-md-6">
```{r}
analysis_data %>%
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= therapy))+
  geom_bar()+
  theme_minimal()+
  labs(title = "Frontline treatment")+
  coord_flip()

# analysis_data %>% 
#   arrange(episodedate) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   group_by(chemotherapy_type) %>% 
#   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>%
#   ggplot(aes(x= chemotherapy_type, y= percent, fill = chemotherapy_type))+
#   geom_bar(stat="identity")+
#   labs(x = "", title = "The first chemotherapy received by a patient Adjuvant for ~70% of the time") +
#   theme_minimal()+
#   guides(fill = FALSE) +
#   coord_flip()
```
</div>

<div class = "col-md-6">
```{r}
analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  group_by(treatment_sequence) %>% 
  summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>%
  ggplot(aes(x= treatment_sequence, y= percent, fill = treatment_sequence))+
  geom_bar(stat="identity")+
  labs(x = "", title = "The treatment_sequence") +
  theme_minimal()+
  guides(fill = FALSE) +
  coord_flip()

# analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
#   ggplot(aes(x= chemo))+
#   geom_bar()+
#   theme_minimal()+
#   coord_flip()
```
</div>
</div>

A quick summary for the RDI repartition
```{r}
analysis_data2 <- analysis_data# %>% filter(linenumber_new == "1")
summary(analysis_data2$mean_RDI_per_drug_chemotype)
```

<div class = "row">
<div class = "col-md-6">
```{r}
analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
  geom_boxplot()+
  ylim(c(-.25, 3.5))+
  geom_jitter(shape=16)+
  labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") +
  theme_minimal()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
  geom_violin()+
  ylim(c(-.25, 3.5))+
  # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
  theme_minimal()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+
  geom_violin()+
  ylim(c(-.25, 3.5))+
  # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
  theme_minimal()+
  coord_flip()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= RDI_grp, fill = drugname))+
  geom_bar()+
  theme_minimal()+
  coord_flip()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= RDI_grp1, fill = drugname))+
  geom_bar()+
  theme_minimal()+
  coord_flip()
```
</div>

<div class = "col-md-6">
```{r}
analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
  # geom_boxplot()+
  ylim(c(-.25, 3.5))+
  geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
  labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") +
  theme_minimal() +
  guides(color=FALSE)

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype, color = mean_RDI_per_drug_chemotype<0.85))+
  geom_violin()+
  ylim(c(-.25, 2.5))+
  scale_colour_discrete(name=NULL, labels=c("<0.85",">=0.85"))+
  # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
  theme_minimal()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+
  geom_violin(aes(color=mean_RDI_per_drug_chemotype<0.85))+
  ylim(c(-.25, 2.5))+
  scale_colour_discrete(name=NULL, labels=c("<0.85",">=0.85"))+
  # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
  theme_minimal()+
  coord_flip()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= RDI_grp, fill = drugname))+
  geom_bar(position = "fill")+
  theme_minimal()+
  coord_flip()

analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
  # filter(linenumber_new == "1") %>% 
  ggplot(aes(x= RDI_grp1, fill = drugname))+
  geom_bar(position = "fill")+
  theme_minimal()+
  coord_flip()
```
</div>
</div>

<!-- ## Extra plots all line together -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- analysis_data %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= therapy))+ -->
<!--   geom_bar()+ -->
<!--   theme_minimal()+ -->
<!--   labs(title = "Frontline treatment")+ -->
<!--   coord_flip() -->

<!-- analysis_data %>% arrange(episodedate) %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!--   filter(exclude != "1") %>%  -->
<!--   group_by(chemotherapy_type) %>%  -->
<!--   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>% -->
<!--   ggplot(aes(x= chemotherapy_type, y= percent, fill = chemotherapy_type))+ -->
<!--   geom_bar(stat="identity")+ -->
<!--   labs(x = "", title = "The first chemotherapy received by a patient Adjuvant for ~70% of the time") + -->
<!--   theme_minimal()+ -->
<!--   guides(fill = FALSE) + -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- analysis_data %>% arrange(episodedate) %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!--   filter(exclude != "1") %>%  -->
<!--   group_by(treatment_sequence) %>%  -->
<!--   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>% -->
<!--   ggplot(aes(x= treatment_sequence, y= percent, fill = treatment_sequence))+ -->
<!--   geom_bar(stat="identity")+ -->
<!--   labs(x = "", title = "The treatment_sequence") + -->
<!--   theme_minimal()+ -->
<!--   guides(fill = FALSE) + -->
<!--   coord_flip() -->

<!-- # analysis_data %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!-- #   ggplot(aes(x= chemo))+ -->
<!-- #   geom_bar()+ -->
<!-- #   theme_minimal()+ -->
<!-- #   coord_flip() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- # What are the RDI distribution for each type of chemo -->
<!-- print("A quick summary for the RDI repartition") -->
<!-- summary(analysis_data$mean_RDI_per_drug_chemotype) -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_boxplot()+ -->
<!--   geom_jitter(shape=16)+ -->
<!--   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") + -->
<!--   theme_minimal() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   # geom_boxplot()+ -->
<!--   ylim(c(-.5, 3.5))+ -->
<!--   geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") + -->
<!--   theme_minimal() + -->
<!--   guides(color=FALSE) -->

<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_violin()+ -->
<!--   ylim(c(-.5, 3.5))+ -->
<!--   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   theme_minimal() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- # What are the RDI distribution for each drug -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_violin(aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   ylim(c(-.5, 5.5))+ -->
<!--   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->

<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= RDI_grp, fill = drugname))+ -->
<!--   geom_bar()+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= RDI_grp, fill = drugname))+ -->
<!--   geom_bar(position = "fill")+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->