### 2021-09-09 Rmd/Report 2021.Rmd
Add more drug cleaning, plots, tables in rmd

---
title: "Flatiron analysis"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: flatly
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 1px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```
<br>

***

```{r library, include=FALSE}
library(drake)
library(tidyverse)
library(data.table)
library(lubridate)
library(ggplot2)
library(gtsummary)
library(survminer)
library(survival)
```

```{r loadd}
loadd(drugs)
creatinine <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/creatinine.rds")
weight <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/weight.rds")
height <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/height.rds")
body_surface_area <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/body_surface_area.rds")
areaUC <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/areaUC.rds")
clinical_data <- read_rds("/Users/colinccm/Documents/GitHub/Peres/Treatment_Flatiron/clinical_data.rds")
```

At this point, after cleaning the clinical data and excluding unfit patients, we have `r nrow(clinical_data)` patients.

# I. Drug episode questions

```{r drugs episode cleanup}
drugs1 <- drugs %>% 
  filter(episodedatasource == "Administrations" & 
           ismaintenancetherapy == "FALSE") %>%
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  # bind with clinical
  inner_join(., clinical_data, #%>% select("patientid", "diagnosisdate", "surgerydate"),
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "months")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "months")) %>% 
  ungroup()

drugs1 %>% filter(days_at_first_drug < 0) %>% distinct(patientid) %>% nrow()
drugs1 %>% filter(days_at_surg < 0) %>% distinct(patientid) %>% nrow()
```

Patients with a date of drug or date of surgery BEFORE diagnosis will be excluded

```{r}
drugs1 <- drugs1 %>% 
  filter(days_at_first_drug >= 0 &
           (is.na(surgerydate) | days_at_surg >= 0)
         )
drugs1 %>% distinct(patientid) %>% nrow()
```

60 days rule between diagnosis and treatment to eliminate patients who didn't get treatment for 60 days which could mean they were treated but it is not recorded 
```{r}
remove_60daysrule <- drugs1 %>% 
  filter((days_at_first_drug > 60 | days_at_surg > 60)) %>% 
  distinct(patientid)
remove_60daysrule <- paste(remove_60daysrule$patientid, collapse = "|")

drugs1 <- drugs1 %>% 
  # remove these patients who 60 days rule
  filter(!str_detect(patientid, remove_60daysrule))

drugs1 %>% distinct(patientid) %>% nrow()
```

Remove linename that doesn't include platin or taxels -> later, need to recode linenumber first
```{r}
# drugs1 <- drugs1 %>% 
#   # 2. Filtering----
#   filter(episodedatasource == "Administrations" & 
#            # ismaintenancetherapy == "FALSE" & 
#            !is.na(amount) &
#            (days_at_drug < 60 | days_at_surg < 60) &
#            
#            days_at_drug >= 0 &
#            (is.na(surgerydate) | days_at_surg >= 0)
#          )

drugs1 %>% 
  filter(linenumber == 1) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linename) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

```{r combine amount same drug one day}
drugs1 <- drugs1 %>% 
  #When want to remove patients who didn't have taxel/platin as 1st line -> remove 260 patients
  # filter(!(linenumber == 1 & !str_detect(linename, "taxel|platin"))) %>% 
  # filter(str_detect(drugname, "taxel|platin") & str_detect(route, "venous|peritoneal")) %>% ############## LATER------------!!!!!
  select(-c(enhancedcohort, episodedatasource, drugcategory, detaileddrugcategory, route,
            diagnosisdate)) %>%

  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, drugname, units, surgerydate, days_at_drug, days_at_surg) %>%
  summarise_at(vars(amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  purrr::keep(~!all(is.na(.))) %>%
  ungroup() %>% 
  mutate(amount_1 = as.numeric(amount_1), amount_2 = as.numeric(amount_2),
         amount_3 = as.numeric(amount_3), amount_4 = as.numeric(amount_4)) %>%
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

  mutate(drugname_type = case_when(
    str_detect(drugname, "platin")       ~ "platin",
    str_detect(drugname, "taxel")        ~ "taxel"
  ))
```

Recode linenumber depending on surgery date
```{r line coding}
therapy_line <- drugs1 %>% 
  # 4. Define Adjuvant, Neoadjuvant, chemo or surgery----
  mutate(chemotherapy_type = case_when(
    # ismaintenancetherapy == "TRUE"    ~ "maintenance",
    episodedate <= surgerydate        ~ "Neoadjuvant", # F63ABD6AEB12C, F4F7B1C5DF24F intraperitoneal Carboplatin
    episodedate > surgerydate         ~ "Adjuvant",
    is.na(surgerydate) &
      !is.na(episodedate)             ~ "Chemotherapy only",
    !is.na(surgerydate) &
      is.na(episodedate)              ~ "Surgery only"
    )) %>% 
  mutate(chemotherapy_type = factor(chemotherapy_type, levels = c("Neoadjuvant", "Adjuvant", "Chemotherapy only", "Surgery only"))) %>%
  group_by(patientid) %>% 
  mutate(had_neo = ifelse(str_detect(chemotherapy_type, "Neo"), "Yes", NA_character_)) %>% 
  fill(had_neo, .direction = "updown") %>% 
  mutate(had_adj = ifelse(str_detect(chemotherapy_type, "Adj"), "Yes", NA_character_)) %>% 
  fill(had_adj, .direction = "updown") %>% 
  mutate(treatment_sequence = case_when(
    had_neo == "Yes" &
      had_adj == "Yes"                           ~ "neo + surg + adj",
    had_neo == "Yes" &
      is.na(had_adj)                             ~ "neo + surg",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "surg + adj",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only"
  )) %>% 
  mutate(therapy = case_when(
    chemotherapy_type == "Neoadjuvant"           ~ "Upfront Chemo",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "Upfront Surgery"
  )) %>% 
  fill(therapy, .direction = "downup") %>% 
  # mutate(therapy1 = case_when(
  #   chemotherapy_type == "Adjuvant"              ~ "Upfront Surgery",
  # )) %>% 
  # fill(therapy1, .direction = "downup") %>% 
  # mutate(therapy = coalesce(therapy, therapy1)) %>% 
  # select( -therapy1) %>%

  # 5.Redefine line number based on the original linenumber variable and the surgery date----
  mutate(linenumber_new = dense_rank(interaction(linenumber, chemotherapy_type))) %>%  # F0000AD999ABF, F1C68CFAC2AF3, FBCF69031DFF8, F0040EA299CF0, FAACDC7205803,  FA019D2071321 no has big jump F95D860690A93 still weird
  select(patientid, linename, linenumber, linenumber_new, drugname, episodedate, surgerydate, chemotherapy_type, everything()) %>% 
  ungroup() %>%

  # 6.Add a 30 day rule to switch linenumber----
  # group_by(patientid, linenumber_new) %>% 
  # arrange(episodedate) %>% 
  # mutate(episode_interval = episodedate - lag(episodedate), episode_interval = ifelse(is.na(episode_interval), 0, episode_interval)) %>%  # FDB58481AFFBB 48 days
  # mutate(jump_line = ifelse(episode_interval > 31, 1, 0)) %>% # FA019D2071321 >31
  # ungroup() %>% 
  # group_by(patientid) %>% 
  # mutate(jump_line = cumsum(jump_line)) %>% # F002DD2953889 fixed even when increase multiple time 
  # mutate(linenumber_new = linenumber_new + jump_line) %>% 
  #F002DD2953889 need to do 37 and 42 F0040EA299CF0, 35 F005602DA4DF0 => do at least 37

  # 7.Calculate cycle for each line----
  mutate(cycle_drugname = str_remove(drugname, " protein-bound")) %>%  # F003F2BEEDB37 need to change drugname pacli = pacli bp
  group_by(patientid, linenumber_new, cycle_drugname) %>% # Wrong line name F001443D8B85C ~~~~~~~~~~~~~~~~~~~~~~~~~~ Fix later QUESTION Dr Chern is ok
  mutate(drugname_count_perline = n()) %>%  # F0000AD999ABF
  ungroup() %>% 
  mutate(across(c("linenumber", "linenumber_new"), as.factor))

# Number of patient with a jump line in the first line # 60 would have
# therapy_line %>% 
#   filter(linenumber == 1 & linenumber_new == 2) %>% 
#   arrange(desc(jump_line)) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(jump_line) %>% 
#   tbl_summary()
```

# Number of patient who switch line because of surgery from line 1 to adjuvant line
```{r line coding}
therapy_line %>% 
  mutate(change_line = case_when(
    linenumber == 1 & 
      linenumber_new == 2 &
      chemotherapy_type == "Adjuvant" &
      treatment_sequence == "neo + surg + adj"     ~ 1,
    TRUE                                           ~ 0
  )) %>% 
  arrange(desc(change_line)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(change_line) %>% 
  tbl_summary()
```

# How many lines patient got as Neo
```{r}
therapy_line %>% 
  filter(chemotherapy_type == "Neoadjuvant") %>% 
  arrange(desc(linenumber_new)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linenumber_new) %>% 
  tbl_summary()

therapy_line %>% 
  filter(chemotherapy_type == "Neoadjuvant" &
           linenumber_new == 2 | linenumber_new == 3 | linenumber_new == 4 | linenumber_new == 5) %>% 
  select(patientid, linename, linenumber, linenumber_new, linestartdate, lineenddate,
         drugname, episodedate, surgerydate, chemotherapy_type) %>%
  arrange(patientid, episodedate)
```

# Remove patients with more than 1 line as Neo -34
```{r}
remove_morethan1_neo <- therapy_line %>% 
  filter(chemotherapy_type == "Neoadjuvant" &
           linenumber_new != 1)
remove_morethan1_neo <- paste(remove_morethan1_neo$patientid, collapse = "|")

therapy_line <- therapy_line %>% 
  # Remove patients with more than 1 line as Neo
  filter(!str_detect(patientid, remove_morethan1_neo))

therapy_line %>% distinct(patientid) %>% nrow()
```

# Remove patients who switch cispl vs carbo # 111 patients Car > Cis F1BF7208F0E96, 34 Cis > Car in neo or adj or chemo
```{r}
# But a lot switch cispl vs carbo # 59 patients Car > Cis F1BF7208F0E96, 32 Cis > Car
therapy_line %>% 
  filter(str_detect(linename, "Car") &
           drugname == "cisplatin") %>%
  distinct(patientid) %>% nrow()

therapy_line %>% 
  filter(str_detect(linename, "Cisplatin") &
           drugname == "carboplatin") %>%
  distinct(patientid) %>% nrow()

remove_switch <- therapy_line %>% 
  filter(
    (str_detect(linename, "Carboplatin") & drugname == "cisplatin") |
  (str_detect(linename, "Cisplatin") & drugname == "carboplatin")
  ) %>%
  distinct(patientid)
remove_switch <- paste(remove_switch$patientid, collapse = "|")

therapy_line <- therapy_line %>% 
  # remove these patients who switched C > C 
  filter(!str_detect(patientid, remove_switch))
therapy_line %>% distinct(patientid) %>% nrow()
```

Should I do the same between paclitaxel and docetaxel? Remember I pacli == pacli protein bound

If a patient had carbo, pacli, carbo, pacli, docetaxel in a Cabo-Pacli line : Could be annoying when I code cycle

In 1 cycle we can have taxel multiple times, is it possible for carbo or cisplatin?
```{r}
# But a lot switch cispl vs carbo # 59 patients Car > Cis F1BF7208F0E96, 32 Cis > Car
therapy_line %>% 
  filter(str_detect(linename, "Pacli") &
           drugname == "docetaxel") %>%
  distinct(patientid) %>% nrow()
therapy_line %>% 
  filter(str_detect(linename, "Doce") &
           drugname == "Paclitaxel") %>%
  distinct(patientid) %>% nrow()
```

# Code cycle
```{r cycle coding}
# Code cycle number
therapy_cycle <- therapy_line %>%

  arrange(patientid, episodedate) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline)) %>%
  mutate(cycle_drugname = case_when(
    cycle_count_perline == drugname_count_perline          ~ drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_drugname) %>%
  mutate(cycle_inc = row_number(cycle_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------

  # + What to do when it is a cycle change F003F2BEEDB37 ,

  # group_by(patientid, linenumber_new, cycle_inc) %>%
  # mutate(cycle_count_1 = row_number(cycle_inc)) %>% 
  # mutate(cycle_c = ifelse(cycle_count_1 == 1, cycle_inc, NA_real_)) %>% 
  arrange(patientid, episodedate, cycle_inc) %>% 
  group_by(patientid, linenumber_new) %>%
  mutate(cycle_increment = cycle_inc) %>% 
  fill(cycle_increment, .direction = "downup") %>%  # make sure I can do up for have one new drug compare to the linename FBB5186A15CD9, FE593EDF5586A
  # cycle date
  group_by(patientid, linenumber_new, cycle_increment) %>% 
  mutate(cycle_start_date = min(episodedate)) %>% 
  # select(patientid, linename, linenumber_new, drugname, episodedate,
  #        cycle_count_perline, cycle_drugname, drugname_count_perline, cycle_increment, cycle_start_date) %>%

  group_by(patientid, linenumber_new) %>% 
  mutate(cycle_interval = cycle_start_date - lag(cycle_start_date), 
         cycle_interval = ifelse(cycle_interval == 0, NA_real_, cycle_interval)) %>%  # FDB58481AFFBB 48 days

  # take for granted that the cycle interval days are the routine cycle for the rest of the line FAACDC7205803
  mutate(n = row_number(drugname)) %>%

  mutate(base_days_between_cycle = ifelse(n==2 , cycle_interval, NA_real_)) %>%
  fill(base_days_between_cycle, .direction = "downup") %>% 
  mutate(skipped_cycle_indays = ifelse(((cycle_interval - base_days_between_cycle) != 0) ,
                                       (cycle_interval - base_days_between_cycle), NA_real_)) %>% # weird FDB58481AFFBB
  # Fix skipped_cycle_indays when only 1 drug FDB58481AFFBB skipped_cycle_indays = NA, no other choice...
  group_by(patientid, linenumber_new) %>%
  mutate(drug_count_perline = n()) %>%
  mutate(skipped_cycle_indays1 = ifelse(
    (cycle_count_perline == drug_count_perline), NA_real_, skipped_cycle_indays 
  )) %>% # F001443D8B85C Pb with when different pattern by cycle => May not be able to have the skipped cycle
  # or need to compare with dose like if 100 should be 7 days and if 200 should be 14 days....
  group_by(patientid) %>% 
  mutate(mean_skipped_cycle_indays = mean(skipped_cycle_indays, na.rm = TRUE)) %>% 
  ungroup()
```

## Bind everything together and calculating target_auc and expected dose

When I bind things together I take the closest date of BSA, AUC, from the cycle date to calculate `target_auc` and `expected_dose`. it can be days in between or more. Do we need a max?
Precision when binding for AUC I bind the closest with corresponding drugname (Pacli AUC is not bind with Carbo drug)
```{r target_auc and expected dose}
Treatment <- 
  # Merge with creatinine
  left_join(therapy_cycle, creatinine, by = "patientid") %>%
  mutate(interval_creat = abs(interval(start= creat_date, end= cycle_start_date)/                      
                          duration(n=1, unit="days"))) %>% 
  arrange(interval_creat) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

  # Merge with height
  left_join(. , height,  by = "patientid") %>%
  mutate(interval_height = abs(interval(start= height_date, end= cycle_start_date)/
           duration(n=1, unit="days"))) %>%
  arrange(interval_height) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
           cycle_start_date, .keep_all = TRUE) %>%

  # Merge with weight
  left_join(. , weight, by = "patientid") %>%
  mutate(interval_weight = abs(interval(start= weight_date, end= cycle_start_date)/
                        duration(n=1, unit="days"))) %>%
  arrange(interval_weight) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
         cycle_start_date, .keep_all = TRUE) %>%

  # Calculate bmi
  mutate(bmi_at_dx = weight_at_dx / (height_at_dx * height_at_dx)) %>%
  mutate(bmi_cat = case_when(
    bmi_at_dx < 25                    ~ "Underweight and normal weight",
    bmi_at_dx >= 25 &
      bmi_at_dx < 30                  ~ "Overweight",
    bmi_at_dx >= 30                   ~ "Obese"
  )) %>%
  mutate(bmi_cat = factor(bmi_cat, levels = c("Underweight and normal weight", "Overweight", "Obese"))) %>%
  # Calculate CrCl
  # inner_join(., clinical_data %>% select(-c("diagnosisdate", "surgerydate")),
  #            by = "patientid") %>% 
  mutate(CrCl = ((140 - ageatdx) * weight)/((72 * creatinine) * 0.85)) %>%
  mutate(CrCl = case_when(
    CrCl > 125                        ~ 125,
    TRUE                              ~ CrCl
  )) %>%

  # Merge with body_surface_area
  left_join(., body_surface_area, by = "patientid") %>%
  mutate(interval_bsa = abs(interval(start= bsa_date, end= cycle_start_date)/
                          duration(n=1, unit="days"))) %>%
  arrange(interval_bsa) %>%
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment,
           cycle_start_date, .keep_all = TRUE) %>%
  # Calculate more bsa value with height and weight using Du Bois formula
  mutate(bsa_du_bois = 0.007184 * ((height*100)^0.725) * (weight^0.425)) %>% # get pretty close results 4269 patients result
  mutate(BSA = coalesce(BSA, bsa_du_bois)) %>%
  mutate(bsa_du_bois = 0.007184 * ((height_at_dx*100)^0.725) * (weight_at_dx^0.425)) %>% # get pretty close results 4269 patients result
  mutate(bsa_at_dx = coalesce(bsa_at_dx, bsa_du_bois)) %>%
  select(-bsa_du_bois) %>%

  # Merge with auc
  left_join(., areaUC, by = c("patientid", "drugname")) %>%
  mutate(interval_auc = abs(interval(start= auc_date, end= cycle_start_date)/ # use episode date to be more precise? No it's not better
                          duration(n=1, unit="days"))) %>% 
  arrange(interval_auc) %>% 
  distinct(patientid, episodedate, drugname, amount, linenumber_new, cycle_count_perline, cycle_increment, 
           cycle_start_date, .keep_all = TRUE) %>%

  # Calculate expected dose
  mutate(expected_dose = case_when(
    drugname == "carboplatin"                 ~ target_auc * (CrCl + 25),
    str_detect(drugname, "taxel|platin")      ~ BSA * target_auc 
    # F19EBBFC60652 Can we fill target_auc? No depends of the type of cycle, or maybe by grouping by cycle type and drug
  )) %>% 
  mutate(expected_dose = case_when(
    expected_dose > 900 &
      target_auc == 6              ~ 900,
    expected_dose > 750 &
      target_auc == 5              ~ 750,
    TRUE                           ~ expected_dose
  )) %>% 
  mutate(first_treatment_date = case_when(
    str_detect(therapy, "Chemo")                      ~ min(episodedate),
    str_detect(therapy, "Surgery")                    ~ surgerydate,
  )) %>% 
  mutate(month_at_os_from_treatment = interval(start = first_treatment_date, end = followupdate)/
                   duration(n=1, units = "months")) %>% 
  arrange(patientid, episodedate)
```

```{r plot interval}
Treatment %>% 
  ggplot(aes(x= interval_creat)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_height)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_weight)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_bsa)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
Treatment %>% 
  ggplot(aes(x= interval_auc)) +
  geom_bar() +
  coord_cartesian(ylim=c(0,50))
```

# Number of patients
```{r}
Treatment %>% distinct(patientid) %>% nrow()
```

Number of line in neo
```{r}
Treatment %>% 
  filter(chemotherapy_type == "Neoadjuvant") %>% 
  arrange(desc(linenumber_new)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linenumber_new) %>% 
  tbl_summary()
```
Number of line in Adj neo + surg + adj
```{r}
Treatment %>% 
  filter(chemotherapy_type == "Adjuvant" & treatment_sequence == "neo + surg + adj") %>% 
  mutate(n = dense_rank(linenumber_new)) %>% 
  arrange(desc(n)) %>%
  distinct(patientid, .keep_all = TRUE) %>%
  select(n) %>% 
  tbl_summary()
```
Number of line in Adj surg + adj
```{r}
Treatment %>% 
  filter(chemotherapy_type == "Adjuvant" & treatment_sequence == "surg + adj") %>% 
  mutate(n = dense_rank(linenumber_new)) %>% 
  arrange(desc(n)) %>%
  distinct(patientid, .keep_all = TRUE) %>%
  select(n) %>% 
  tbl_summary()
```
Number of line in neo
```{r}
Treatment %>% 
  filter(chemotherapy_type == "Chemotherapy only") %>% 
  arrange(desc(linenumber_new)) %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  select(linenumber_new) %>% 
  tbl_summary()
```

# Final dataset
```{r analysis_data}
analysis_data <- Treatment %>% 
  # distinct(patientid, chemotherapy_type, drugname, .keep_all = TRUE) %>% 
  filter(exclude != "1")
```

After excluding exclude == "1" patients, we have `r analysis_data %>% distinct(patientid) %>% nrow()` patients.

```{r}
library(naniar)
analysis_data %>%
  distinct(patientid, .keep_all = TRUE) %>% 
  select(bmi_at_dx, weight_at_dx, height_at_dx, bsa_at_dx) %>% 
  gg_miss_upset()

analysis_data %>% select(creatinine, CrCl, target_auc, expected_dose) %>% 
  gg_miss_upset()

is.na(analysis_data$CrCl) %>% sum()
is.na(analysis_data$amount) %>% sum()
is.na(analysis_data$target_auc) %>% sum()

# analysis_data %>% filter(is.na(amount)) %>%
#     distinct(patientid, .keep_all = TRUE) %>% nrow()
# 
# analysis_data %>% filter(is.na(target_auc) | is.na(expected_dose)) %>%
#     distinct(patientid, .keep_all = TRUE) %>% nrow()
```
If we remove patients who have at least 1 missing amount administrered, we lose `r analysis_data %>% filter(is.na(amount)) %>% distinct(patientid, .keep_all = TRUE) %>% nrow()` patients.

If we remove patients who are missing at least 1 target_auc or 1 expected dose we lose `r analysis_data %>% filter(is.na(target_auc) | is.na(expected_dose)) %>% distinct(patientid, .keep_all = TRUE) %>% nrow()` patients.

# CALCULATE RDI

```{r}
Frontline <- Treatment %>% 
  # filter(therapy == "Upfront Chemo" | therapy == "Chemotherapy only") %>% 
  # group_by(patientid, linenumber_new, cycle_increment, drugname) %>% # ! -----------------------Change drugname if combine taxel....
  # mutate(amount_cycle = sum(amount)) %>% 
  # select(patientid, linenumber_new, cycle_increment, drugname, amount, amount_cycle, 
  #        target_auc, expected_dose, everything()) %>% # F01E4E6DD63C5, F1FA39C41A087 ----------- No we have xpected dose for each episode
  # ungroup() %>% 
  # distinct(patientid, linenumber_new, cycle_increment, drugname, .keep_all = TRUE) %>%

  # Create a linenumber restating at 1 for adj to be able to calculate RDI for first line of Adjuvant
    group_by(patientid, chemotherapy_type) %>% 
  mutate(linenumber_adj = dense_rank(linenumber_new)) %>%

  # Calculate 1 RDI per drug per day, for whole Neoadjuvant and for first line of Adjuvant
  mutate(relative_dose_intensity = case_when(
    chemotherapy_type == "Adjuvant" &
      linenumber_adj == 1                  ~ amount/expected_dose,
    chemotherapy_type == "Neoadjuvant"     ~ amount/expected_dose
  )) %>%
  # select(patientid, linenumber_new, chemotherapy_type, drugname_type, drugname,
  #        relative_dose_intensity, linenumber_adj) %>%

  # Average RDI by taxel or platin by line (will combine pacli-doce and cis-carbo-oxalo)
  group_by(patientid, chemotherapy_type, drugname_type) %>%
  mutate(mean_RDI_per_chemotype_drug = case_when(
    chemotherapy_type == "Adjuvant" &
      linenumber_adj == 1                  ~ round(mean(relative_dose_intensity, na.rm = TRUE), 3),
    chemotherapy_type == "Neoadjuvant"     ~ round(mean(relative_dose_intensity, na.rm = TRUE), 3)
  )) %>% 
  ungroup() %>%

  # Then let's see if we need to average that for the whole Neo BUT 
  # For now average rdi over whole neo, average rdi for the first line over adjuvant
  # when plot meed 1 rdi per drugname_type per chemotherapy_type aka distinct(patientid, chemotherapy_type, drugname_type)

  # group_by(patientid, chemotherapy_type, drugname_type) %>% 
  # mutate(RDI_neo = case_when(
  #   # chemotherapy_type == "Adjuvant"       ~ first(mean_RDI_per_drug_line), # F002B429BE6C6
  #   chemotherapy_type == "Adjuvant" &
  #     linenumber_adj == 1                  ~ mean(mean_RDI_per_drug_line),
  #   chemotherapy_type == "Neoadjuvant"     ~ mean(mean_RDI_per_drug_line) # F0040EA299CF0
  # )) %>%

  # mutate(RDI_adj = case_when(
  #   chemotherapy_type == "Adjuvant" &
  #     linenumber_adj == 1                  ~ mean(mean_RDI_per_drug_line), # F002B429BE6C6
  #   # chemotherapy_type == "Neoadjuvant"     ~ mean(mean_RDI_per_drug_line) # F0040EA299CF0
  # )) %>% 
  # mutate(RDI = coalesce(RDI_neo, RDI_adj))

  # group_by(patientid, chemotherapy_type, drugname) %>% # ------------------------ Or should I do by cycle, then avg again
  # mutate(mean_RDI_per_drug_chemotype = mean(relative_dose_intensity)) %>% 
  # ungroup() %>% 
  mutate(RDI_grp = as.factor(findInterval(mean_RDI_per_chemotype_drug, c(0.85) ))) %>% 
  mutate(RDI_grp = factor(RDI_grp, 
                           levels = c("1", "0"), 
                           labels = c("RDI >= 0.85", "RDI < 0.85"))) %>% 
  mutate(RDI_grp1 = as.factor(findInterval(mean_RDI_per_chemotype_drug, c(0.75, 0.85, 1, 1.5, 2) ))) %>% 
  mutate(RDI_grp1 = factor(RDI_grp1, 
                          levels = c("3", "0", "1","2","4", "5"), 
                          labels = c("0.85 >= RDI < 1", "0 < RDI < 0.75", "0.75 <= RDI < 0.85", "1 <= RDI < 1.5", 
                                     "1.5 <= RDI < 2", "RDI >= 2"))) %>% 
  # mutate(delay_incare_dx_to_treat = case_when(
  #   str_detect(therapy, "Chemo")     ~ (interval(start = diagnosisdate, end = episodedate)/
  #     duration(n=1, units = "days")),
  #   str_detect(therapy, "Surgery")   ~ (interval(start = diagnosisdate, end = surgerydate)/
  #     duration(n=1, units = "days"))
  # )) %>% 
  # group_by(patientid) %>% 
  # mutate(first_treatment_date = case_when(
  #   treatment_sequence == "Chemotherapy only"             ~ min(episodedate),
  #   str_detect(treatment_sequence, "neo")                 ~ min(episodedate),
  #   treatment_sequence == "surg + adj"                    ~ surgerydate,
  #   treatment_sequence == "Surgery only"                  ~ surgerydate
  # )) %>% 
  group_by(patientid) %>% 
  mutate(first_treatment_date = case_when(
    str_detect(therapy, "Chemo")                      ~ min(episodedate),
    str_detect(therapy, "Surgery")                    ~ surgerydate,
  )) %>% 
  mutate(delay_incare_dx_to_treat = interval(start = diagnosisdate, end = first_treatment_date)/
                                          duration(n=1, units = "days")) %>% 
  mutate(month_at_os_from_treatment = interval(start = first_treatment_date, end = followupdate)/
                   duration(n=1, units = "months")) %>% 
  ungroup() %>% 
  # %>% 
  # select(patientid, linenumber_new, drugname, surgerydate, episodedate, treatment_sequence, first_treatment_date, month_at_os_from_treatment)
  mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/
           duration(n=1, units = "days"), 
         thirty_days_exclusion = ifelse(thirty_days_exclusion <= 30, "Exclude", NA_character_))

```

# Analysis

## 1.Clinical summary table

<span style="color:green">**Table 1 Overall summary demographic/clinical data**</span>
```{r Table 1}

# How should I filter??????????????????///
analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
  select(c(raceeth, 
           agecat, "ageatdx",

           bmi_at_dx, bmi_cat, 
           weight_at_dx, height_at_dx, bsa_at_dx,

           "histology", "stagecat",

           "brca1_germline", "brca2_germline", "brca_germline", "brca1_tumor",
           "brca2_tumor", "brca_tumor", "brca1_unknown", "brca2_unknown", "brca_unknown")) %>% 
  tbl_summary() %>% bold_labels()
```

<span style="color:green">**Table 2 Summary demographic/clinical data by race**</span>
```{r Table 2}
analysis_data %>% 
  select(c(raceeth, "vital",
           agecat, "ageatdx",

           "bmi_at_dx", bmi_cat, "BSA",

           "histology", "groupstage", "tstage", "stagecat",

           "brca1_germline", "brca2_germline", "brca_germline", "brca1_tumor",
           "brca2_tumor", "brca_tumor", "brca1_unknown", "brca2_unknown", "brca_unknown",

           "hrd_tissue", "gis_tissue")) %>% 
  tbl_summary(by = raceeth) %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
```
<br>

## 2.Plots Clinical
<div class = "row">
<div class = "col-md-6">
```{r}
analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= bmi_cat))+
  geom_bar()+
  theme_minimal()+
  labs(title = "BMI category")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= histology))+
  geom_bar()+
  theme_minimal()+
  labs(title = "Histology")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= stagecat))+
  geom_bar()+
  theme_minimal()+
  labs(title = "Stage")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= agecat))+
  geom_bar()+
  theme_minimal()+
  labs(title = "Age cat")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= brca_germline))+
  geom_bar()+
  theme_minimal()+
  labs(title = "BRCA gernline")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= brca_tumor))+
  geom_bar()+
  theme_minimal()+
  labs(title = "BRCA tumor")+
  coord_flip()
```
</div>

<div class = "col-md-6">
```{r}
analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= raceeth, fill = bmi_cat))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "BMI category")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= histology, fill = raceeth))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "Histology")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= stagecat, fill = raceeth))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "Stage")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= agecat, fill = raceeth))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "Age cat")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= brca_germline, fill = raceeth))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "BRCA germline")+
  coord_flip()

analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= brca_tumor, fill = raceeth))+
  geom_bar(position = "fill")+
  theme_minimal()+
  labs(title = "BRCA tumor")+
  coord_flip()
```
</div>
</div>

<br>

## 3.Survivals by clinical

<span style="color:green">**Survival plots from date of diagnosis on the left and from date of treatment on the right**</span>

<div class = "row">
<div class = "col-md-6">
```{r clinical from dx, fig.height = 7}
# OS by race
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ race, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by raceeth
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ raceeth, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by histology
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ histology, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by tstage
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ stagecat, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by bmi_cat
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ bmi_cat, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca1_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca1_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca2_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca_tumor
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca1_g
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca1_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca2_g
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca2_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca_g
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ brca_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
```
</div>

<div class = "col-md-6">
```{r clinical from first treatment, fig.height = 7}
# OS by race
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ race, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by raceeth
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ raceeth, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by histology
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ histology, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by tstage
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ stagecat, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by bmi_cat
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ bmi_cat, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca1_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca1_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca2_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca2_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca_tumor
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca_tumor, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
# OS by brca1_g
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca1_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca2_g
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca2_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by brca_g
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ brca_germline, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
```
</div>
</div>

<br>
<br>

***
# II.Analysis Treatment

## 1.Treatment summary table

<span style="color:green">**Table 3 Summary treatment data by race**</span>

```{r Table 3}
# tbl1 <- 
  analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
  filter(exclude != "1") %>% 
  select(c(raceeth, #delay_incare_dx_to_treat, #mean_skipped_cycle_indays,
           "issurgery", "extentofdebulking", "resdz", "debulking", "diff_surg_dx",
           had_neo, had_adj,
           "chemotherapy_type", "therapy", "treatment_sequence"#, "mean_RDI_per_drug_chemotype", "RDI_grp"
           )) %>% 
  tbl_summary(by = raceeth) %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
#   filter(exclude != "1") %>% 
#   select(c(raceeth, delay_incare_dx_to_treat, #mean_skipped_cycle_indays,
#            "issurgery", "extentofdebulking", "resdz", "debulking", "diff_surg_dx",
#            "chemotherapy_type", "therapy", "treatment_sequence"#, "mean_RDI_per_drug_chemotype", "RDI_grp"
#            )) %>% 
#   tbl_summary(by = raceeth, missing = "no") %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("with unknown", "without unknown"))
```
<br>

## 2.Plots Treatment
<div class = "row">
<div class = "col-md-6">
```{r}
analysis_data %>%
  distinct(patientid, .keep_all = TRUE) %>% 
  ggplot(aes(x= therapy))+
  geom_bar()+
  theme_minimal()+
  labs(title = "Frontline treatment")+
  coord_flip()

# analysis_data %>% 
#   arrange(episodedate) %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   group_by(chemotherapy_type) %>% 
#   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>%
#   ggplot(aes(x= chemotherapy_type, y= percent, fill = chemotherapy_type))+
#   geom_bar(stat="identity")+
#   labs(x = "", title = "The first chemotherapy received by a patient Adjuvant for ~70% of the time") +
#   theme_minimal()+
#   guides(fill = FALSE) +
#   coord_flip()
```
</div>

<div class = "col-md-6">
```{r}
analysis_data %>% 
  distinct(patientid, .keep_all = TRUE) %>% 
  # group_by(treatment_sequence) %>% 
  # summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>%
  ggplot(aes(x= treatment_sequence,
             # y= percent, 
             fill = treatment_sequence))+
  geom_bar(
    # stat="identity"
    )+
  labs(x = "", title = "The treatment_sequence") +
  theme_minimal()+
  guides(fill = FALSE) +
  coord_flip()

# analysis_data %>% distinct(patientid, .keep_all = TRUE) %>% 
#   ggplot(aes(x= chemo))+
#   geom_bar()+
#   theme_minimal()+
#   coord_flip()
```
</div>
</div>

<br>

## 3.Survivals by treatment

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ issurgery, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by extentofdebulking
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ extentofdebulking, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by residualdiseasestatus
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ residualdiseasestatus, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by chemotherapy_type
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ chemotherapy_type, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by treatment_sequence
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ treatment_sequence, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by therapy
ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ therapy, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of diagnosis", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ issurgery, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by extentofdebulking
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ extentofdebulking, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by residualdiseasestatus
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ residualdiseasestatus, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by chemotherapy_type
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ chemotherapy_type, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))

# OS by treatment_sequence
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ treatment_sequence, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(nrow = 3))

# OS by therapy
ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ therapy, data=analysis_data %>% 
                     distinct(patientid, .keep_all = TRUE) %>% 
                     filter(exclude != "1")),
           title = "OS from date of from first treatment", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           # Add risk table
           risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
           risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),

           legend.title = "", xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
```
</div>
</div>

<br>
<br>

***
# III.RDI

## 1.RDI summary table

<span style="color:green">**Table 4 Summary RDI data by race and bmi**</span>

```{r Table 4}

# Frontline %>% group_by(chemotherapy_type, drugname) %>% summarize(average_RDI =mean(mean_RDI_per_drug_chemotype, na.rm = TRUE)) 
# 
# # Treatment & demo
# tbl1 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth,  "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               label = list(mean_RDI_per_drug_chemotype ~ "carb/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat,  "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               label = list(mean_RDI_per_drug_chemotype ~ "carb/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl3 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl1 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               label = list(mean_RDI_per_drug_chemotype ~ "carb/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               label = list(mean_RDI_per_drug_chemotype ~ "carb/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl4 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl1 <- analysis_data %>% filter(str_detect(drugname, "cisplatin") & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               label = list(mean_RDI_per_drug_chemotype ~ "cisplatin/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(str_detect(drugname, "cisplatin") & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               label = list(mean_RDI_per_drug_chemotype ~ "cisplatin/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl5 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl1 <- analysis_data %>% filter(str_detect(drugname, "cisplatin") & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               type = list(mean_RDI_per_drug_chemotype ~ 'continuous'),
#               label = list(mean_RDI_per_drug_chemotype ~ "cisplatin/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(str_detect(drugname, "cisplatin") & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               type = list(mean_RDI_per_drug_chemotype ~ 'continuous'),
#               label = list(mean_RDI_per_drug_chemotype ~ "cisplatin/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl6 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl_stack(list(tbl4, tbl3, tbl5, tbl6), quiet = TRUE)
# 
# 
# 
# tbl1 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               label = list(mean_RDI_per_drug_chemotype ~ "taxel/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Adjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               label = list(mean_RDI_per_drug_chemotype ~ "taxel/adjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl5 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl1 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(raceeth, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = raceeth,
#               label = list(mean_RDI_per_drug_chemotype ~ "taxel/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl2 <- analysis_data %>% filter(str_detect(drugname, "taxel") & chemotherapy_type == "Neoadjuvant") %>% 
#   filter(exclude != "1") %>% 
#   distinct(patientid, .keep_all = TRUE) %>% 
#   select(c(bmi_cat, "mean_RDI_per_drug_chemotype", RDI_grp)) %>% 
#   tbl_summary(by = bmi_cat,
#               label = list(mean_RDI_per_drug_chemotype ~ "taxel/neoadjuvant \nRDI")) %>%
#   bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
# tbl6 <- 
#   tbl_merge(list(tbl1, tbl2), tab_spanner = c("Race/Ethnicity", "BMI"))
# 
# tbl_stack(list(tbl5, tbl6), quiet = TRUE)
```
<br>

## 2.Plots RDI
For RDI boxplot and violin, it is calculated for neoadjuvant and adjuvant when a patient got both (need to find another way of plotting to do by drug - maybe a wrap).

<div class = "row">
<div class = "col-md-6">
```{r}
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
#   geom_boxplot()+
#   ylim(c(-.25, 3.5))+
#   geom_jitter(shape=16)+
#   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") +
#   theme_minimal()
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
#   geom_violin()+
#   ylim(c(-.25, 3.5))+
#   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   theme_minimal()
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+
#   geom_violin()+
#   ylim(c(-.25, 3.5))+
#   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   theme_minimal()+
#   coord_flip()
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= RDI_grp, fill = drugname))+
#   geom_bar()+
#   theme_minimal()+
#   coord_flip()

```
</div>

<div class = "col-md-6">
```{r}
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+
#   # geom_boxplot()+
#   ylim(c(-.25, 3.5))+
#   geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") +
#   theme_minimal() +
#   guides(color=FALSE)
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype, color = mean_RDI_per_drug_chemotype<0.85))+
#   geom_violin()+
#   ylim(c(-.25, 2.5))+
#   scale_colour_discrete(name=NULL, labels=c(">=0.85", "<0.85"))+
#   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   theme_minimal()
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+
#   geom_violin(aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   ylim(c(-.25, 2.5))+
#   scale_colour_discrete(name=NULL, labels=c(">=0.85", "<0.85"))+
#   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   theme_minimal()+
#   coord_flip()
# 
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= RDI_grp, fill = drugname))+
#   geom_bar(position = "fill")+
#   theme_minimal()+
#   coord_flip()
#

```
</div>
</div>

```{r}
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype, alpha = mean_RDI_per_drug_chemotype<0.85))+
#   geom_violin(aes(fill = drugname))+
#   ylim(c(-.25, 2.5))+
#   scale_colour_discrete(name=NULL, labels=c(">=0.85", "<0.85"))+
#   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+
#   theme_minimal()
```

A quick summary for the RDI repartition
```{r}
# analysis_data2 <- analysis_data# %>% filter(linenumber_new == "1")
# summary(analysis_data2$mean_RDI_per_drug_chemotype)
```

<div class = "row">
<div class = "col-md-6">
```{r}
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= RDI_grp1, fill = drugname))+
#   geom_bar()+
#   theme_minimal()+
#   coord_flip()
```
</div>

<div class = "col-md-6">
```{r}
# analysis_data %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>%
#   # filter(linenumber_new == "1") %>% 
#   ggplot(aes(x= RDI_grp1, fill = drugname))+
#   geom_bar(position = "fill")+
#   theme_minimal()+
#   coord_flip()
```
</div>
</div>

<br>

## 3.Survivals by RDI

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
# OS by RDI_grp
# ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      distinct(patientid, .keep_all = TRUE) %>% 
#                      filter(exclude != "1")),
#            title = "OS from diagnosis Overall RDI", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
# 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for platin
# ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "carboplatin")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from diagnosis carboplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "cisplatin")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from diagnosis cisplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for taxel
# ggsurvplot(survfit(Surv(month_at_os, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "taxel")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from diagnosis all taxel", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
# # OS by RDI_grp
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      distinct(patientid, .keep_all = TRUE) %>% 
#                      filter(exclude != "1")),
#            title = "OS from first treatment overall RDI", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for platin
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "carboplatin")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment carboplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "cisplatin")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment cisplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for taxel
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "taxel")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from from first treatment all taxel", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
```
</div>
</div>

<br>

## Survivals RDI by chemotherapy type
**On the left are Neoadjuvant, on the right are Adjuvant**

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
# # OS by RDI_grp for platin
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "carboplatin"),
#                             str_detect(chemotherapy_type, "Neoadjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment carboplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "cisplatin"),
#                             str_detect(chemotherapy_type, "Neoadjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment cisplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for taxel
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "taxel"),
#                             str_detect(chemotherapy_type, "Neoadjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from from first treatment all taxel", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
# # OS by RDI_grp for platin
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "carboplatin"),
#                             str_detect(chemotherapy_type, "Adjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment carboplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "cisplatin"),
#                             str_detect(chemotherapy_type, "Adjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from first treatment cisplatin", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
# 
# # OS by RDI_grp for taxel
# ggsurvplot(survfit(Surv(month_at_os_from_treatment, vitalstatus) ~ RDI_grp, data=Frontline %>% 
#                      filter((exclude != "1") & str_detect(drugname, "taxel"),
#                             str_detect(chemotherapy_type, "Adjuvant")) %>% 
#                      
#                      distinct(patientid, .keep_all = TRUE)), 
#            title = "OS from from first treatment all taxel", 
#            font.main = c(24, "bold", "black"), 
#            font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
#            font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
#            size = 1.5,
#            
#            pval=TRUE, 
#            # legend.labs=c(""),
#            # palette = c("black", "green","blue", "red"), 
#            # Add risk table
#            risk.table = "abs_pct", risk.table.title = "Risk table (number(%))",
#            risk.table.y.text = FALSE, risk.table.fontsize = 3, tables.height = 0.3,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            
#            legend.title = "", xlab = "Time (in months)" 
# ) + guides(colour = guide_legend(ncol = 2))
```
</div>
</div>

<br>
<br>

***
# IV.Hazard Ratio

<span style="color:green">**Table 5 HR**</span>  
`coxph(Surv(time = month_at_os, event = vitalstatus)~variables`
```{r Table 5 HR}
# # tbl1 <- analysis_data %>% 
# #   select(bmi_cat) %>% 
# #   tbl_summary(#by = vitalstatus,
# #               statistic = all_categorical() ~ "{analysis_data %>% filter(vital == 'Dead') %>% NROW()} death ({n})"
# #               )
# # tbl2 <- coxph(Surv(time = month_at_os,
# #                      event = vitalstatus)~bmi_cat,
# #                 data = analysis_data) %>%
# #   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# # 
# # tbl_merge(list(tbl1, tbl2))
# # 
# # coxph(Surv(time = month_at_os, 
# #                      event = vitalstatus)~RDI_grp, 
# #                 data = analysis_data) %>% 
# #   tbl_uvregression(y = vitalstatus, exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# 
# analysis_data %>% select(vitalstatus, bmi, bmi_cat, raceeth) %>% 
# tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = analysis_data$month_at_os, event = analysis_data$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   bold_labels() %>% italicize_levels() #%>%  modify_spanning_header(all_stat_cols() ~ "**Treatment Received**")
# 
# tbl1 <- analysis_data %>% 
#   filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant")
# tbl1 <- tbl1 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl1$month_at_os, event = tbl1$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl2 <- analysis_data %>% 
#   filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant")
# tbl2 <- tbl2 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl2$month_at_os, event = tbl2$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl3 <- analysis_data %>% 
#   filter(drugname == "carboplatin" & str_detect(chemotherapy_type, "Chemo"))
# tbl3 <- tbl3 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl3$month_at_os, event = tbl3$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl_stack(list(tbl1, tbl2, tbl3),
#           group_header = c("carboplatin/Neoadjuvant", "carboplatin/Adjuvant", "carboplatin/Chemo only")) %>% 
#   bold_labels() %>%
#   italicize_levels()
# # Cisplatin
# tbl4 <- analysis_data %>% 
#   filter(str_detect(drugname,"cisplatin") & chemotherapy_type == "Neoadjuvant")
# tbl4 <- tbl4 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl4$month_at_os, event = tbl4$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl5 <- analysis_data %>% 
#   filter(str_detect(drugname,"cisplatin") & chemotherapy_type == "Adjuvant")
# tbl5 <- tbl5 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl5$month_at_os, event = tbl5$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl6 <- analysis_data %>% 
#   filter(str_detect(drugname,"cisplatin") & str_detect(chemotherapy_type, "Chemo"))
# tbl6 <- tbl6 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl6$month_at_os, event = tbl6$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl_stack(list(tbl4, tbl5, tbl6),
#           group_header = c("cisplatin/Neoadjuvant", "cisplatin/Adjuvant", "cisplatin/Chemo only")) %>% 
#   bold_labels() %>%
#   italicize_levels()
# 
# 
# tbl7 <- analysis_data %>% 
#   filter(str_detect(drugname,"paclitaxel") & chemotherapy_type == "Neoadjuvant")
# tbl7 <- tbl7 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl7$month_at_os, event = tbl7$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl8 <- analysis_data %>% 
#   filter(str_detect(drugname,"paclitaxel") & chemotherapy_type == "Adjuvant")
# tbl8 <- tbl8 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl8$month_at_os, event = tbl8$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   modify_header(update = list(label ~ "**paclitaxel/Adjuvant**"))
# tbl9 <- analysis_data %>% 
#   filter(str_detect(drugname,"paclitaxel") & str_detect(chemotherapy_type, "Chemo"))
# tbl9 <- tbl9 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl9$month_at_os, event = tbl9$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl_stack(list(tbl7, tbl8, tbl9),
#           group_header = c("paclitaxel/Neoadjuvant", "paclitaxel/Adjuvant", "paclitaxel/Chemo only")) %>% 
#   bold_labels() %>%
#   italicize_levels()
# 
# tbl10 <- analysis_data %>% 
#   filter(str_detect(drugname, "taxel") & chemotherapy_type == "Neoadjuvant")
# tbl10 <- tbl10 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl10$month_at_os, event = tbl10$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl11 <- analysis_data %>% 
#   filter(str_detect(drugname, "taxel") & chemotherapy_type == "Adjuvant")
# tbl11 <- tbl11 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl11$month_at_os, event = tbl11$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   modify_header(update = list(label ~ "**paclitaxel/Adjuvant**"))
# tbl12 <- analysis_data %>% 
#   filter(str_detect(drugname,"taxel") & str_detect(chemotherapy_type, "Chemo"))
# tbl12 <- tbl12 %>% 
#   select(vitalstatus, RDI_grp, RDI_grp1) %>% tbl_uvregression(method = survival::coxph, 
#                  y = (Surv(time = tbl12$month_at_os, event = tbl12$vitalstatus)),
#                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent()
# tbl_stack(list(tbl10, tbl11, tbl12),
#           group_header = c("taxel/Neoadjuvant", "taxel/Adjuvant", "taxel/Chemo only")) %>% 
#   bold_labels() %>%
#   italicize_levels()
```
<br>
<br>

***
# V.Regression
## By clinical, demo
`glm(vitalstatus ~ bmi + bmi_cat + RDI_grp + raceeth, data = analysis_data, family = binomial)`
```{r}
# Regression vital status prediction
model <- glm(vitalstatus ~ bmi_at_dx + bmi_cat + #RDI_grp + 
               raceeth, data = analysis_data, family = binomial)
tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

`glm(vitalstatus ~ histology + tstage + brca1_tumor + brca2_tumor + issurgery + chemotherapy_type, data = analysis_data, family = binomial)`
```{r}
model <- glm(vitalstatus ~ 
               histology + tstage + brca1_tumor + brca2_tumor + 
             issurgery +
             chemotherapy_type, 
             data = analysis_data, family = binomial)
tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

`glm(RDI_grp ~ agecat + bmi_cat + raceeth + stagecat + histology + chemotherapy_type, data = analysis_data, family = binomial)`
<!-- !!!! Prediction of RDI all drug together!!!! -->
```{r}
# Regression RDI prediction
# model <- glm(RDI_grp ~ agecat + bmi_cat + raceeth + stagecat + histology + chemotherapy_type, data = analysis_data, family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
```

<!-- ### By drug -->
<!-- !!! But sometimes we had a switch between cisplatin/carboplatin -->
```{r}
# #  for carbo
# model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
#                filter(drugname == "carboplatin" & chemotherapy_type == "Adjuvant"), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl3 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
#                filter(drugname == "carboplatin" & chemotherapy_type == "Neoadjuvant"), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl4 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# # tbl5 <- 
#   tbl_merge(list(tbl3, tbl4), tab_spanner = c("**carboplatin/Adjuvant**", "**carboplatin/Neoadjuvant**"))
#   
# #  for Cisplatim
# model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
#                filter(drugname == "cisplatin" & chemotherapy_type == "Adjuvant"), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl3 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# model <- glm(RDI_grp ~ agecat + bmi_cat, data = analysis_data %>% 
#                filter(drugname == "cisplatin" & chemotherapy_type == "Neoadjuvant"), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl4 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# # tbl5 <- 
#   tbl_merge(list(tbl3, tbl4), tab_spanner = c("**cisplatin/Adjuvant**", "**cisplatin/Neoadjuvant**"))
# 
# #  for taxel
# model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
#                filter(str_detect(drugname,"taxel") & (chemotherapy_type == "Adjuvant")), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl3 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# model <- glm(RDI_grp ~ agecat + raceeth + stagecat + histology + bmi_cat, data = analysis_data %>% 
#                filter(str_detect(drugname,"taxel") & (chemotherapy_type == "Neoadjuvant")), family = binomial)
# tbl1 <- tbl_regression(model) %>% bold_p(t = .05)
# tbl2 <- tbl_regression(model) %>% bold_p(t = .05) %>% bold_p(t = .05)
# tbl4 <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Estimate**", "**Exp**"))
# 
# # tbl6 <- 
#   tbl_merge(list(tbl4, tbl3), tab_spanner = c("**taxel/Adjuvant**", "**taxel/Neoadjuvant**"))
```
<br>
<br>

***

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->

<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->

<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- ## Extra plots all line together -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- analysis_data %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= therapy))+ -->
<!--   geom_bar()+ -->
<!--   theme_minimal()+ -->
<!--   labs(title = "Frontline treatment")+ -->
<!--   coord_flip() -->

<!-- analysis_data %>% arrange(episodedate) %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!--   filter(exclude != "1") %>%  -->
<!--   group_by(chemotherapy_type) %>%  -->
<!--   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>% -->
<!--   ggplot(aes(x= chemotherapy_type, y= percent, fill = chemotherapy_type))+ -->
<!--   geom_bar(stat="identity")+ -->
<!--   labs(x = "", title = "The first chemotherapy received by a patient Adjuvant for ~70% of the time") + -->
<!--   theme_minimal()+ -->
<!--   guides(fill = FALSE) + -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- analysis_data %>% arrange(episodedate) %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!--   filter(exclude != "1") %>%  -->
<!--   group_by(treatment_sequence) %>%  -->
<!--   summarize(count = n()) %>% mutate(percent=(count/sum(count)*100)) %>% -->
<!--   ggplot(aes(x= treatment_sequence, y= percent, fill = treatment_sequence))+ -->
<!--   geom_bar(stat="identity")+ -->
<!--   labs(x = "", title = "The treatment_sequence") + -->
<!--   theme_minimal()+ -->
<!--   guides(fill = FALSE) + -->
<!--   coord_flip() -->

<!-- # analysis_data %>% distinct(patientid, .keep_all = TRUE) %>%  -->
<!-- #   ggplot(aes(x= chemo))+ -->
<!-- #   geom_bar()+ -->
<!-- #   theme_minimal()+ -->
<!-- #   coord_flip() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- # What are the RDI distribution for each type of chemo -->
<!-- print("A quick summary for the RDI repartition") -->
<!-- summary(analysis_data$mean_RDI_per_drug_chemotype) -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_boxplot()+ -->
<!--   geom_jitter(shape=16)+ -->
<!--   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") + -->
<!--   theme_minimal() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   # geom_boxplot()+ -->
<!--   ylim(c(-.5, 3.5))+ -->
<!--   geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   labs(x = "", title = "Warning 1 dot = RDI for 1 drugname per patient") + -->
<!--   theme_minimal() + -->
<!--   guides(color=FALSE) -->

<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= chemotherapy_type, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_violin()+ -->
<!--   ylim(c(-.5, 3.5))+ -->
<!--   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   theme_minimal() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- # What are the RDI distribution for each drug -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= drugname, y= mean_RDI_per_drug_chemotype))+ -->
<!--   geom_violin(aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   ylim(c(-.5, 5.5))+ -->
<!--   # geom_jitter(shape=16, aes(color=mean_RDI_per_drug_chemotype<0.85))+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->

<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= RDI_grp, fill = drugname))+ -->
<!--   geom_bar()+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r} -->
<!-- Frontline %>% filter(!is.na(mean_RDI_per_drug_chemotype)) %>% -->
<!--   filter(exclude != "1") %>%  -->
<!--   ggplot(aes(x= RDI_grp, fill = drugname))+ -->
<!--   geom_bar(position = "fill")+ -->
<!--   theme_minimal()+ -->
<!--   coord_flip() -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->